\textbf{\sc Towards Deliberative Control of Autonomous Robots}

### 2012-02-19 arch.tex
further updates to arch

\subsubsection{Intertwining synchronization and execution}
\label{sec:arch:intertwine}

One key aspect of the \rx framework is intertwined plan execution. As
seen above, it is synchronization which drives planning, so the
co-mingling between these two processes is critical to maintaining
state within a robot immersed in a dynamic environment. Also important
is that \rx maintains 
% A core aspect of this reactor is that it takes advantage of the
% specific sequencing enforced by the \rx architecture along with the
% fact that only 
one $plan$ structure in the reactor and shared between synchronization
and planning. This allows these two processes to not only be
interleaved but how they can impact each other via plan manipulation.

The most obvious case is how synchronization allow propagate of {\em
  external} observations as a new tick occurs. In our presentation of 
{\em synchronization}, we isolated the problem by stating
that we will assume at this point that the reactor has no future goal
or more accurately no plan to enforce. By doing so we
were able to develop the way we augment the \eu solver to be able
to propagate {\em external} observations in order to identify the
reactor {\em internal} state for the same tick. This allowed us 
to focus on synchronization as a state evaluation process that can
be resolved within the \eu framework with few extensions. Similarly
we presented deliberation at this fairly high level not necessarily
discussing how the disruptive nature of synchronization in this
process could impact the way this planning could evolve.

We first analyze the case during which synchronization occurs after
planning has completed (\ie the last step of deliberation results in a
plan that is considered complete for the current look-ahead). As
synchronization starts with a plan % produced this will gave a frame for
% the search process. Therefore 
this deliberation will not only be a model-derived estimation uniquely
based on {\em external} observations but will also include reactor
intentions specified in the available partial plan. Consider for
example the Shopping Agent model described where the \texttt{Agent} is
an {\em external} timeline. As the reactor adds the goal to
\texttt{Own} milk and the current {\em external} timeline indicates
that the robot was \texttt{At(home)}, it produces a plan as shown in
Fig. \ref{fig:shop:exec0}. The {\em external} part of the plan --
namely the tokens \texttt{Going(home, superMarket)} and
\texttt{At(SuperMarket)} -- are installed as goals for the reactor
managing this state variable {\em internally}.

@@

Consider that at the next tick $\tau' = \tau+1$ the reactor managing
agent location is able to change the location state variable to
\texttt{Going}. \rx informs our reactor of this new observation which
is added in the plan structure as a fact -- starting exactly at
$\tau'$ and lasting for an unknown duration which is at least 1
tick. As this new observation is integrated in the plan through
synchronization, the solver can identify the similarity between this
observation and the next token in the existing plan. Instead of
initiating an exhaustive search of all possible implications, the
synchronization process starts its execution with the assumption that
these new observations are consistent with the currently maintained
plan. As a consequence, at the end of synchronization the plan
depicted in Fig. \ref{fig:shop:exec1} is likely. This is a result of
propagation of information provided by the new observation in the
existing plan while restricting the start time of \texttt{Going} and
propagating temporal constraints, making \texttt{Go} to start {\em
  exactly} at $\tau'$. In doing so, the \eu solver's focus on the
execution frontier ($\tau'$ in Fig. \ref{fig:shop:exec1}) takes
advantage of the existing plan to evaluate a solution that assumes
that the observation is consistent with what was decided by the
reactor.

@@

This choice is directed not only by heuristic choices made in the
solver but also enforced by the heuristic order of the decision point
resolution choices for the {\em current  state flaw} presented in
Proposition \ref{prop:csf:resolve}. Indeed, as we stated then the two
first choices tend to privilege a conservative view in respect to the
existing plan. It attempts first to extend the previous observation we
produced for this {\em internal} timeline in order to avoid to commit
to early in the plan advance. If the formaer failed it will then
attempt to start the next exisiting token in the plan which alow the
reactor to advance its plan execution. Both of these two initial
choices atetempt to resolve the plan with the least amount of
perturbation of the plan. They just constrain further the start and
end time of the exisiting tokens which were provided by the previous 
deliberations steps. This means that in a nominal situation where the
plan this reactor produced is properly executed, the synchronization
remains limited to restrict time points in the plan and progate their
impact. Most of the heavy dutty work of creating new tokens in the
plan and schedule them -- both being much more difficult problems 
to resolve -- have been managed during deliberation steps. In that
sense we see that most of the time synchronization is limited to very
few node in \eu search space with often no backtrack required. 
Fig. \ref{fig:shop:exec} shows how the existing plan can support the
synchronization process. In that case, the observation \texttt{Going}
occuring at $\tau'$ was fully compatible with the one already present
in our plan, by updating the plan and progating the equality
constraints, the resolution of the {\em current state flaw} for the
{\em agent} timeline became obvious as the token \texttt{Go} from 
the plan presented itself as a natural solution.

The synchronization may become slightly nore costfull when we need to
identify a new token to be injected by the plan (the 3rd choice in
Proposition \ref{prop:csf:resolve}) as it needs to explore on tis own
all the possible token types this timeline accept as value. This is
the reason this choice is evaluated last in our flaw resolution in
order to avoid such complex action during synchronization unless 
necessary. In that sense,  and thanks to the carefull choice on our
synchronization heurisitic we see that the plan produced during \rx
deliberation steps helps sthe synchronization process to be more
efficient.

@@

Synchronization also supports the tracking of plan execution. It does
so by specifying token parameters such as time points that are
propagated through the plan in order to identify if the plan remains
valid -- as in our example -- or that it cannot be executed as it
is. The later is often identified by the synchronization failing to
find a consistent solution. For example should the \texttt{Going}
observation received be to the \texttt{hardwareStore}, this would
break the current partial plan in the reactor as the robot cannot
possible be going simultaneously to $2$ different locations; see
Fig. \ref{fig:shop:relax1}. As synchronization is not possible in such
a situation the strategy followed is to relax all decisions made by
previous plan steps. This is done by keeping all past observations in
both {\em external} and {\em internal} state variables as well as the
existing {\em internal} goals (\texttt{Own(Milk)} in this example)
received by this reactor. The resulting partial plan is shown in
Fig. \ref{fig:shop:relax2}. In this process the reactor informs the
agent that all {\em external} goals it requested for the future are no
longer valid which results in turn, in timeline owners not having to
maintain such goals.

@@

As the plan is no longer conflicted, synchronization can resume and
find a solution described above. Newly pending goals would be enforced
at the next \texttt{step} to resume deliberation on this relaxed
partial plan. By using synchronization the reactor was not only able
to identify that its plan was not consistent but able to recover from
it by offering a new incomplete partial plan to be resolved in the
next deliberation \texttt{steps}. % This on its own offer the core
% functionality one should except to close the loop between planning and
% execution in the sense that 
Synchronization plays the role of what is typically termed as the {\em
  Executive} in other architectures \cite{gat98, alami:1998p820,
  mus98, williams03, Nesnas:2003do}.

In \rx however the interaction between planning and execution does not
stop at this level. The overall scheduling of both {\em
  synchronization} and {\em planning} is such that the synchronization
can occur in between any planning step. Not only is synchronization
influenced by the outcome, but it also influences
planning. Specifically when synchronization occurs before a complete
plan is found, it will inject new facts ({\em external} observations
and resulting {\em internal} state) in the plan database available for
inference when the next planning \texttt{step} resumes. On this next
step the \eu solver used for planning will have these new facts
included in the partial plan which can introduce new flaws in the plan
while potetnially having reolved previously exisiting flaw. The
deliberation process is therefore perturbed externally by the {\em
  synchronization} and consequently fully informed of the world state
evolution as it is planning.  This specific aspect contrasts with
other more classical approaches we have seen, which avoid 
perturbing the planner during its search as it is not compatible 
with the ``off-line planning'' assumption.

By our design choices -- both at the architecture level and the
integration of the \eu framework for embedded deliberation -- we are
able to implement an architecture that avoids this critical off-line
planning limitation while remaining within a formal framework that
allows for deliberation and model-based agent control. We believe that
this tight integration of planning and execution allows the system to
be more acute vis-a-vis the environment allowing informed
decision-making and a tighter reaction loop. This is possible while
integrating information despite a long latency for the reactors, as
long as one can ensure that synchronization of all reactors and
deliberation steps can be done within a time that is resonably small
in comparison to the agent tick duration.

>>>>

\subsubsection{Interleaving Synchronization and Execution}
\label{sec:arch:intertwine}

A key aspect of the \rx framework is interleaved plan execution. As
noted above, it is synchronization which drives planning, so the
co-mingling between these two processes is critical to maintaining
state within a robot especially when it is immersed in a dynamic
environment. Further, it is important that \rx maintain
% A core aspect of this reactor is that it takes advantage of the
% specific sequencing enforced by the \rx architecture along with the
% fact that only 
one $plan$ structure shared between synchronization and planning. This
allows these two processes to not only be interleaved for agent
responsiveness, but ensures a viable mechanism to communicate.

% but also impact
% each other via plan manipulation.

As noted earlier, synchronization propagates {\em
  external} observations when a new tick occurs. Earlier, we isolated
the case when a reactor has no future goal or no plan to enforce. By
doing so we were able to develop the way we augment the \eu solver to
propagate {\em external} observations to identify {\em
  internal} state. This allowed us to focus on synchronization as a
state evaluation process that can be resolved within the \eu framework
with few extensions. Similarly we presented deliberation as an
abstraction without discussing the disruptive nature of
synchronization and how it impacts the planning process. We first
analyze the case when synchronization occurs after planning has
completed (\ie the last step of deliberation results in a plan that is
considered complete for the current look-ahead). As synchronization
starts with a plan, this deliberation will not only be a model-derived
estimation uniquely based on {\em external} observations but will also
include reactor intentions specified in the available partial plan.

Consider for example the Shopping Agent model where the \texttt{Agent}
is an {\em external} timeline. As the reactor adds a goal to
\texttt{Own} milk and the current {\em external} timeline indicates
that the robot was \texttt{At(home)}, it produces a plan as shown in
Fig. \ref{fig:shop:exec0}. The {\em external} part of the plan --
namely the tokens \texttt{Going(home, superMarket)} and
\texttt{At(SuperMarket)} -- are installed as goals for the reactor
managing this state variable internally.

@@

Consider that at the next tick $\tau' = \tau+1$ the reactor managing
agent location is able to change the location state variable to
\texttt{Going}. \rx informs the reactor of this observation which is
added to the partial plan as a fact -- starting exactly at $\tau'$ and
lasting for an unknown duration which is at least one tick. As this
new observation is integrated in the plan through synchronization, the
solver can identify the similarity between this observation and the
next token in the existing plan. Instead of initiating an exhaustive
search of all possible implications, the synchronization process
starts its execution with the assumption that these new observations
are consistent with the currently maintained plan. As a consequence,
at the end of synchronization the plan depicted in Fig.
\ref{fig:shop:exec1} is likely. This is a result of propagation of
information provided by the new observation in the existing plan while
restricting the start time of \texttt{Going} and propagating temporal
constraints, making \texttt{Go} to start {\em exactly} at $\tau'$. In
doing so, the \eu solver's focus on the execution frontier ($\tau'$ in
Fig. \ref{fig:shop:exec1}) takes advantage of the existing plan to
evaluate a solution that assumes that the observation is consistent
with what was decided by the reactor.

@@

This choice is directed not only by heuristic choices made in the
solver but also enforced by the heuristic order of the decision point
resolution choices for the {\em current state flaw} shown in
Proposition \ref{prop:csf:resolve}. As noted, the first two choices
are biased towards a conservative view with respect to the existing
plan. The heuristics first attempt to extend the previous observation
for this {\em internal} timeline to avoid early commitment. If this
fails \rx attempts to start the next existing token in the plan to
advance plan execution. Both choices attempt to resolve state with the
least amount of plan perturbation by constraining further the start
and end time of existing tokens generated by previous deliberations
steps. By doing so, in a nominal situation, synchronization remains
limited to restrict time points in the plan and propagates their
impact. Most of the computational effort of creating new tokens in the
plan and scheduling them onto timelines are managed during
deliberation steps. Consequently, most of the time synchronization is
limited to very few nodes in the \eu search space often requiring no
backtracking. Fig. \ref{fig:shop:exec} shows how the existing plan can
support the synchronization process. The observation \texttt{Going}
occuring at $\tau'$ is fully compatible with the token already present
in our plan. By updating the plan and progating equality constraints,
the resolution of the {\em current state flaw} for the {\em Agent}
timeline becomes obvious as a solution, as the token \texttt{Go} from
the plan presents itself.

Synchronization may become computationally more expensive when we need
to identify a new token to be injected into the plan (the third choice
in Proposition \ref{prop:csf:resolve}), as it needs to explore the
outcome of inserting all possible token types the timeline can accept.
This is the reason this choice is evaluated last in our flaw
resolution strategy.

% In that sense, and thanks to the carefull choice on our
% synchronization heurisitic we see that the plan produced during \rx
% deliberation steps helps sthe synchronization process to be more
% efficient.

@@

Synchronization also supports the tracking of plan execution. It does
so by specifying token parameters such as time points that are
propagated throughout the plan to identify plan validity.
%  -- as in our example -- or that it cannot be executed as it is. 
When invalid, identification occurs by synchronization failing to find
a consistent solution. For example should the \texttt{Going}
observation received be to the \texttt{hardwareStore}, this would
break the partial plan in the reactor as the robot cannot possible be
going simultaneously to two different locations; see Fig.
\ref{fig:shop:relax1}. As synchronization is not possible in such a
situation the strategy followed is to relax all decisions made by
previous plan steps. This is done by keeping all past observations in
both {\em external} and {\em internal} state variables as well as the
existing {\em internal} goals (\texttt{Own(Milk)} in this example)
received by the reactor. The resulting partial plan is shown in Fig.
\ref{fig:shop:relax2}. In this process the reactor informs the agent
that all {\em external} goals it requested for the future are no
longer valid, allowing timeline owners to divest maintaining such
goals.

@@

When the plan is no longer conflicted, synchronization can resume and
find a solution as described above. Newly pending goals would be
enforced at the next \texttt{step} to resume deliberation on this
relaxed partial plan. By using synchronization the reactor is not only
able to identify that the plan is consistent but able to recover from
it by offering a new incomplete partial plan to be resolved in the
next deliberation \texttt{steps}. % This on its own offer the core
% functionality one should except to close the loop between planning and
% execution in the sense that 
Synchronization therefore, plays the role of what is typically termed
as the {\em Executive} in other architectures \cite{gat98,
  alami:1998p820, mus98, williams03, Nesnas:2003do}.

In \rx however the interaction between planning and execution does not
stop at this level. The overall scheduling of both {\em
  synchronization} and {\em planning} is such that the synchronization
can occur in between any planning step. Not only is synchronization
influenced by the outcome, but it also influences planning.
Specifically when synchronization occurs before a complete plan is
found, it will inject new facts ({\em external} observations and
resulting {\em internal} state) in the plan database available for
inference when the next planning \texttt{step} resumes. On this next
step the \eu solver used for planning will have these new facts
included in the partial plan which can introduce new flaws in the plan
while potentially having resolved previously existing flaws. The
deliberation process is therefore perturbed externally by {\em
  synchronization} and consequently fully informed of the world state
evolution as it is planning. This contrasts with other classical
approaches, which avoid perturbing the planner during its search as it
is not compatible with the ``off-line planning'' assumption, noted in
Section \ref{sec:arch:europa}.

By design choices made in the \rx framework -- both at the
architectural level and the integration of the \eu framework for
embedded deliberation -- we are able to implement a system that avoids
this critical off-line planning limitation while enforcing a formal
framework that allows for deliberation and model-based agent control.
This tight integration of planning and execution allows the system to
be more responsive to environmental change allowing informed
decision-making within a tighter reaction loop and despite reactor
latency.

% As long as one can ensure that synchronization of all
% reactors and deliberation steps can be done within a time that is
% resonably small in comparison to the agent tick duration.
